Towards a Complete Model of Mutation Selection Balance in Humans

### 2023-05-06 manuscript/manuscript.tex
ms updates

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not have an
effect on fitness or are deleterious. Negative selection works to eliminate
such deleterious mutations from the population, which is essential to
maintaining the integrity of information in the genome to maintain a lineage.
Mutations under negative selection are kept at low frequencies (Haldane?),
contributing to human traits (XXX) and diseases. As with most selection in the
genome, this process perturbs the allele frequencies of neighboring linked
variation. This reduces genetic variability around conserved segments and
creates large-scale spatial patterns in genetic diversity. These patterns are
mediated by the deleterious mutation rate, the strength of selection in
conserved segments, and the spatial distribution of recombination rates and
conserved segments. Since genome-wide recombination maps and putatively
conserved segments are available for many species, researchers have combined
these with population genetic theory to estimate the mutation rate and
selection coefficients of deleterious mutations, as well as the overall
reduction in genetic variability due to the negative selection process.

@@

Similar to previous work, our method characterizes the negative selection
process through its indirect effects on levels of neighboring pairwise
diversity at putatively neutral sites. Essential to this approach are accurate
theoretic models for how a selection process alters pairwise coalescence times
at sites some recombination distance away. The predominant model for how
negative (or purifying) selection reduces pairwise coalescence times at linked
sites is the classic background selection model
\parencite{Charlesworth1993-gb}. This simple and elegant model assumes new
deleterious mutations enter a region of $L$-basepairs at a rate of $\mu$ per
basepair per generation with a selection coefficient $s$ in heterozygotes that
is sufficiently deleterious that it is destined to loss. Under this setup, the
number of segregating deleterious mutations in the region reaches a stationary
Poisson distribution with rate $\nicefrac{U}{s}$ ($U=\mu L$), which is
precisely the equilibrium number under a deterministic mutation-selection
balance models \parencite{Kimura1966-bk,Crow1970-wj}. Modifications of this
basic model that allow for recombination between the segment under negative
selection and a focal neutral site \parencite{Nordborg1996-nq,Hudson1994-oh}
have been the basis of statistical methods that fit genome-wide patterns of
diversity to theoretic expectations \parencite{McVicker2009-ax,Hudson1995-xc}.
Furthermore, jointly incorporating both hitchhiking and background selection
have allowed for model comparison approaches that can estimate the relative
contributions of these two modes of selection
\parencite{Elyashiv2016-vt,Murphy2022-sj}.

These linked selection models predict how selection increases the rate
of coalescence at a neighboring neutral site some recombination distance away.

In general, these methods characterize selection processes by fitting the
relationship between observed pairwise diversity and expected levels according
to a theoretic model of linked selection. These linked selection models predict
how selection impacts the coalescence rate at some nearby site, which is
mediated by the strength of selection (e.g. the rate of new selected

For negative selection, the background
selection model predicts the expected pairwise coalescence time at some focal
neutral site near a region under mutation-selection balance. The increase in
coalescence rate

The impact of selection on linked neutral sites is mediated by the rate at
which selected mutations enter the population, the strength of selection, and
the recombination distance between

Both
hitchhiking and background selection models approximate the increased rate of
coalescence near sites under positive and negative selection as a reduction in
effective population size. Hitchhiking models

typically imagine a sweeping
beneficial allele creating

feature of linked selection theory is that the increased rate of coalescence at
sites neighboring regions under either positive or negative selection

Both positive
and negative selection increase the rate of coalescence at neighboring sites,
which both hitchhiking (XXX) and background selection (XXX) models

>>>>

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not affect
fitness or are deleterious. Negative selection works to eliminate such
deleterious mutations from the population, which is essential to maintaining
the integrity of information in the genome to maintain a lineage. Mutations
under negative selection are kept at low frequencies (Haldane?), contributing
to human traits (XXX) and diseases. As with most selection in the genome, this
process perturbs the allele frequencies of neighboring linked variation. This
reduces genetic variability around conserved segments and creates large-scale
spatial patterns in genetic diversity. These patterns are mediated by the
deleterious mutation rate, the strength of selection in conserved segments, and
the spatial distribution of recombination rates and conserved segments. Since
genome-wide recombination maps and putatively conserved segments are available
for many species, researchers have combined these with population genetic
theory to estimate the mutation rate and selection coefficients of deleterious
mutations, as well as the overall reduction in genetic variability due to the
negative selection process.

@@

Similar to previous work, our method characterizes the negative selection
process through its indirect effects on neighboring levels of pairwise
diversity. While pairwise diversity is a less-informative statistic than the
full site frequency spectrum, it is far easier to derive models of how
selection alters expected pairwise coalescence times at linked sites than to
describe the entire genealogy. Statistical methods then fit observed diversity
to the expectation of a particular theoretic model of the linked selection
process. Consequently, statistical inferences about selection may be incorrect
when the underlying theory inaccurately predicts average pairwise coalescence
times.

Our work is motivated by a well-known limitation of classic background
selection theory: it is only accurate for strongly deleterious mutations. Given
that statistical methods rely on background selection theory to estimate the
mutation rate and distribution of fitness effects of deleterious mutations, it
is important to characterize and remedy these shortcomings of the model.
However, this is

It is well-known that the approximations that underpin classic background
selection theory
\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-pt}
break down when selection against deleterious mutations is weak, and the
reduction in diversity predicted by this model is incorrect
\parencite{McVean2000-bt,Good2013-lp,Gordo2002-dr}. Before introducing how th

describe how selection alters pairwise coalescence times rather than full
genealogy (since $\E[\pi] = 2\mu \E(T_2)$). Statistical methods that fit the
relationship between observed diversity to an expectation provided by a
particular theoretic model of the linked selection process.

Our work considers a well-known
shortcoming of the classic background selection model: that

The classic theoretic model of how negative or purifying selection changes the
mean coalescence time at linked neutral sites is the background selection model
\parencite{Charlesworth1993-gb}. In the simplest case without recombination,
this model assumes that deleterious mutations enter a segment.

classic background selection model are only applicable for strongly deleterious
mutations that have no probability of fixation
\parencite{Good2014-yz,Gordo2002-dr,Charlesworth2013-kl}.

new deleterious mutations enter a region of $L$-basepairs at a rate of $\mu$
per basepair per generation with a selection coefficient $s$ in heterozygotes
that is sufficiently deleterious that it is destined to loss. Under this setup,
the number of segregating deleterious mutations in the region reaches a
stationary Poisson distribution with rate $\nicefrac{U}{s}$ ($U=\mu L$), which
is precisely the equilibrium number under a deterministic mutation-selection
balance models \parencite{Kimura1966-bk,Crow1970-wj}. Since any linear carrying
a deleterious mutation is destined to extinction, coalescence events can only
occur in the $N \exp(-\nicefrac{U}{s})$ individuals carrying zero deleterious
mutations. Working back through the ancestors of two lineages carrying $i$ and
$j$ deleterious mutations, both

Modifications of this basic model that allow for recombination between the
segment under negative selection and a focal neutral site
\parencite{Nordborg1996-nq,Hudson1994-oh} have been the basis of statistical
methods that fit genome-wide patterns of diversity to theoretic expectations
\parencite{McVicker2009-ax,Hudson1995-xc}. Furthermore, methods that
incorporate both hitchhiking and background selection allow for estimates of
the relative contributions of these two modes of selection and model
comparisions \parencite{Elyashiv2016-vt,Murphy2022-sj}.

However, the predicted reduction in diversity under the classic background
selection model is inaccurate whenever selection is weak enough that
deleterious mutations may rise in frequency in the population and possibly
reach fixation. This is a well-known problem (XXX), that is mathematically
difficult because the

\subsection*{DFE Estimates}
 - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
